Analysis of Security-Cost Trade-off of Fully Homomorphic Encryption Schemes

### 2013-06-10 Report/article.tex
Introduction modified

\documentclass{acm_proc_article-sp}
\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax

\usepackage{listings}
\usepackage{algorithm2e}
\begin{document}
\title{Analysis of Security-Cost Trade-off of Fully Homomorphic Encryption Schemes}

@@

\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.
\maketitle
\begin{abstract}
We present a study on variants of fully homomorphic schemes and the trade-off cost/security of basic operations on bits, integers, strings... 
\end{abstract}

\keywords{ Fully Homomorphic Encryption, Security, Cost} % NOT required for Proceedings

\section{Introduction(1 page)}
The homomorphic encryption allows few operations on encrypted data. Gentry was the first to establidh a fully homomorphic encryption scheme based on ideal lattices which allows operations on encrypted data
 but with a high cost. Since then many researches have been made trying to improve the schemes to establish more operations on encrypted data with less cost and better secutity in order to do computations on remote machines. 
In this article we study the implementation of basic classes of operations and different aspects of 
complexity-security, and we discuss the results of experimental measures based on several parameters.\\
The Gentry FHE protocol consists on ``somewhat homomorphic'' scheme, ``squashing'' and ``bootstrapping''. (explain later). 
There are variants of FHE schemes more optimized than Gentry's scheme, so we use in the experiments some open source implementations of these schemes.
\section{Fully homomorphic scheme(1 page)}
A general structure of the scheme.
\subsection{Two Variants of Fully Homomorphic Encryption Scheme}
A brief description of the two FHE schemes with their algorithmic complexity of the different stages in the scheme. 
-> Smart Vercauteren
-> BGV
Discuss the security of these schemes.

\begin{itemize}
\item Smart-Vercauteren scheme:\\
The scheme has 3 parameters: ($N, \nu, \mu$ ). The somewhat homomorphic scheme uses 5 algorithms: (\texttt{KeyGen}, \texttt{Encrypt}, \texttt{Decrypt}, \texttt{Add}, \texttt{Mult}}).
\begin{itemize}
\item \texttt{KeyGen}:
\begin{itemize}
\item $F(x)= x^{n}+1$
\item do
\begin{itemize}
\item $S(x)=_{R}(B_{\infty , N}(\nu/2)$
\item $G(x)=1+2.S(x)$
\item $p= resultant(G(x),F(x))$
\end{itemize}
until $p$ is prime
\item $D(x)=gcd(G(x),F(x))$ over $\mathbb{F}_p[x]$
\item $\alpha$: the unique root of $D(x)$
\item apply \texttt{fmpz\_poly\_xgcd} to obtain $Z(x)$ such that $Z(x).G(x)=p$ mod $F(x)$
\item $B=z_0$ mod $2p$
\item $PK = (p, \alpha)$ and $Sk = (p , B)$
\end{itemize}
\item \texttt{Encrypt}($m \in \{0,1\} , PK$):
\begin{itemize}
\item $R(x)=_{R}(B_{\infty , N}(\mu/2)$
\item $C(x)=m+2.R(x)$
\item $c=C(\alpha) mod p$
\item return $c$
\end{itemize}
\item \texttt{Decrypt}($c, SK$):
\begin{itemize}
\item $m= (c - round(c.B/p)) mod 2$
\item return $m$
\end{itemize}
\item \texttt{Add}($c_1$, $c_2$, $PK$):
\begin{itemize}
\item $c_3=c_1+c_2 mod p$
\item return $c_3$
\end{itemize}
\item \texttt{Mult}($c_1$, $c_2$, PK):
\begin{itemize}
\item $c_3=c_1.c_2 mod p$
\item return $c_3$
\end{itemize}
\end{itemize}
\item BGV:
\end{itemize}
\subsection{Considered Implementations}

@@

The benchmark functions and problems on which the implementations were evaluated are categorized into three classes : functions operating on bits, functions operating directly on integers or block of bits and functions where branching is required i.e \texttt{if-then-else} condition is evaluated.

Functions operating on bits provide information on the tradeoff for calculations with bits, while when operating directly on integers, the tradeoff is obtained on size of integers. we note that the first two categories of functions can be evaluated using straight-line-programs. For others we choose evaluation of \texttt{if-then-else} condition on encrypted data  wich woukd make the evlaluation circuit longer than compared to the same evaluation on unencrypted data. The functions evaluated are :

\begin{itemize}

\item XOR of bits
\item Sum of :

	\begin{itemize}

		\item  bounded integers

		\item unbounded integers

 	\end{itemize}
\item Majority bit

\item sorting with the following methods :

	\begin{itemize}

		\item Insertion sort

		\item Odd-Even Merge sort

		\item Bitonic sort

	\end{itemize}

\end{itemize}

We include the program used to evalute the above functons and their cost in terms of \texttt{XOR} and \texttt{AND} gates.

@@

Sorting is tested on a sorting network where comparator gates are used. Each comparator is designed using the following algorithm.

@@

Each comparator gate uses $n$ \texttt{AND} gates, $n$ \texttt{OR} gates and $n$ \texttt{NOT} gates to find the larger of the two bit sequences and then to regenerate the maximum and the minuimum $4n$ \texttt{AND} gates $2n$ \texttt{NOT} and \texttt{OR} gates i.e. $6n$ \texttt{AND} and \texttt{XOR} gates are required.

\textbf{Insertion sort :} $n(n-1)/2$ comparator gates are used for insertion sort.

\textbf{Odd-Even Merge sort :} The following program is used to perform Odd-Even merge sort.

@@

\begin{figure}
\begin{lstlisting}[caption=Odd-Even Merge Sort]

 OddEvenMergeSort(int lo, int n)
if (n>1)
        {
            int m=n/2;
            oddEvenMergeSort(lo, m);
            oddEvenMergeSort(lo+m, m);
            oddEvenMerge(lo, n, 1);
        }
oddEvenMerge(int lo, int n, int r)
    {
        int m=r*2;
        if (m<n)
        {
	    //even subsequence ;		
            oddEvenMerge(lo, n, m); 
            //odd subsequence ;
	    oddEvenMerge(lo+r, n, m);    
            for (int i=lo+r; i+r<lo+n; i+=m)
                compare(i, i+r);
        }
        else
            compare(lo, lo+r);
    }

\end{lstlisting}

\end{figure}

\textbf{Bitonic Sort}
\begin{figure}
\begin{lstlisting}[caption=Bitonic Sort]

sortup( int m, int n) {//from m to m+n
    if (n==1) return;
    sortup(m,n/2);
    sortdown(m+n/2,n/2);
    mergeup(m,n/2);
}
 sortdown(int m, int n) {//from m to m+n
    if (n==1) return;
    sortup(m,n/2);
    sortdown(m+n/2,n/2);
    mergedown(m,n/2);
}

mergeup(int m, int n) {
    if (n==0) return;

    for (i=0 : n) {
         compare(m+i,m+i+n);
    }
    mergeup(m,n/2);
    mergeup(m+n,n/2);
}
mergedown( int m,  int n) {
    if (n==0) return;
    for (i=0 : n) {
        compare(m+i,m+i+n);
    }
    mergedown(m,n/2);
    mergedown(m+n,n/2);
}
\end{lstlisting}
\caption{Bitonic Search}
\end{figure}
Measurements on varying the parameters or the plaintext space. Evaluation of program dependent properties.

@@

Anlayze the results previously obtained and its impact on security and complexity. Comparisions of security-cost tradeoff with other schemes like RSA.
\section{Conclusion}
Conclusion\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
\bibliographystyle{abbrv}
\bibliography{}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\balancecolumns
% That's all folks!
\end{document}

>>>>

\documentclass{acm_proc_article-sp}
\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage{hyperref}
\usepackage[all]{hypcap}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{algorithm2e}
\hypersetup{
  bookmarks=true,         % show bookmarks bar?
  unicode=false,          % non-Latin characters in Acrobat’s bookmarks
  pdftoolbar=true,        % show Acrobat’s toolbar?
  pdfmenubar=true,        % show Acrobat’s menu?
  pdffitwindow=false,     % window fit to page when opened
  pdfstartview={FitH},    % fits the width of the page to the window
  pdftitle={My title},    % title
  pdfauthor={Author},     % author
  pdfsubject={Subject},   % subject of the document
  pdfcreator={Creator},   % creator of the document
  pdfproducer={Producer}, % producer of the document
  pdfkeywords={keyword1} {key2} {key3}, % list of keywords
  pdfnewwindow=true,      % links in new window
  colorlinks=true,       % false: boxed links; true: colored links
  linkcolor=cyan,          % color of internal links (change box color with linkbordercolor)
  citecolor=green,        % color of links to bibliography
  filecolor=magenta,      % color of file links
  urlcolor=cyan           % color of external links
}

\usepackage{listings}
\usepackage{algorithm2e}
\usepackage{verbatim}
\begin{document}
\title{Analysis of Security-Cost Trade-off of Fully Homomorphic Encryption Schemes}

@@

\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.
\maketitle
\begin{abstract}
This paper presents an analysis of security-cost trade-off for homomorphic encryption schemes. Analysis is based on the measurements taken on the evaluation of certain functions on two  variants of fully homomorphic encryption schemes. These functions either operate on bits, on integers (of bounded and unbounded size)  or involve \texttt{if-then-else} branching. 
\end{abstract}

\keywords{ Fully Homomorphic Encryption, Security, Cost.} % NOT required for Proceedings

\section{Introduction(1 page)}

The notion of \textit{fully homomorphic encryption scheme}, orginally called a \textit{privacy homomorphism} was introduced by Rivest et al. \cite{rivest78} shortly after the invention of the RSA by Rivest, Shamir and Adleman \cite{Rivest78amethod}.  Basic RSA is a multiplicative homomorphic encryption scheme -- i.e. given the public parameters $pk:=(e, N)$ and two messages $m_0, m_1$, their encryption $c_0, c_1$ verify $c_0c_1=(m_0m_1)^e \ \textrm {mod}\ N$. With a fully homomorphic encryption scheme : a scheme $\mathcal{E}$ with an efficient algorithm $Eval_\mathcal{E}$ that evaluates any circuit $C$ on any ciphertexts $c_i \leftarrow Encrypt(pk, m_i)$; one can arbitrarily compute any function on encrypted data without the decryption key. However prior to \cite{homenc}, we did not have a viable construction.

As an application, such schemes can be used to query data on a remote machine. A user prepares his encrypted data $c_0, c_1, \ldots, c_{n-1}$  and a description of the function(or a query) i.e. a circuit $C$ he wants to evaluate and sends them to the remote machine. The machine transforms the circuit into $C^{'}$ so that it can evaluate the same function but on encrypted data, performs the operation and returns the encrypted result to the user. The user eventually decrypts the result.

The homomorphic encryption scheme proposed by Gentry \cite{homenc} is based on ideal lattices which allows operations on encrypted data
but with a high cost depending on the depth of the circuit representing the $Eval$ function. Since then, researchers have improved and proposed other schemes such as \cite{cryptoeprint:2009:571} and \cite{cryptoeprint:2011:277} which decrease the circuit complexity. In this article we expermentally evaluate implementation of these two schemes and analyze the cost-security trade-off for $Eval$ varying from \texttt{XOR} of the bits of an $n$ bit integer to sum of bounded and unbounded integers to sorting of $n$ integers. Three different sorting algorithms are tested.

In this article we study the implementation of basic classes of operations and different aspects of 
complexity-security, and we discuss the results of experimental measures based on several parameters.\\
The Gentry FHE protocol consists on ``somewhat homomorphic'' scheme, ``squashing'' and ``bootstrapping''. (explain later). 
There are variants of FHE schemes more optimized than Gentry's scheme, so we use in the experiments some open source implementations of these schemes.
\section{Fully homomorphic scheme(1 page)}
A general structure of the scheme.
\subsection{Two Variants of Fully Homomorphic Encryption Scheme}
A brief description of the two FHE schemes with their algorithmic complexity of the different stages in the scheme. 
-> Smart Vercauteren
-> BGV
Discuss the security of these schemes.

\begin{itemize}
\item Smart-Vercauteren scheme:\\
The scheme \cite{cryptoeprint:2009:571} has 3 parameters: ($N, \nu, \mu$ ). The somewhat homomorphic scheme uses 5 algorithms: (\texttt{KeyGen}, \texttt{Encrypt}, \texttt{Decrypt}, \texttt{Add}, \texttt{Mult}).
\begin{itemize}
\item \texttt{KeyGen}:
\begin{itemize}
\item $F(x)= x^{n}+1$
\item do
\begin{itemize}
\item $S(x)=_{R}(B_{\infty , N}(\nu/2)$
\item $G(x)=1+2.S(x)$
\item $p= resultant(G(x),F(x))$
\end{itemize}
until $p$ is prime
\item $D(x)=gcd(G(x),F(x))$ over $\mathbb{F}_p[x]$
\item $\alpha$: the unique root of $D(x)$
\item apply \texttt{fmpz\_poly\_xgcd} to obtain $Z(x)$ such that $Z(x).G(x)=p$ mod $F(x)$
\item $B=z_0$ mod $2p$
\item $PK = (p, \alpha)$ and $Sk = (p , B)$
\end{itemize}
\item \texttt{Encrypt}($m \in \{0,1\} , PK$):
\begin{itemize}
\item $R(x)=_{R}(B_{\infty , N}(\mu/2)$
\item $C(x)=m+2.R(x)$
\item $c=C(\alpha) mod p$
\item return $c$
\end{itemize}
\item \texttt{Decrypt}($c, SK$):
\begin{itemize}
\item $m= (c - round(c.B/p)) mod 2$
\item return $m$
\end{itemize}
\item \texttt{Add}($c_1$, $c_2$, $PK$):
\begin{itemize}
\item $c_3=c_1+c_2 mod p$
\item return $c_3$
\end{itemize}
\item \texttt{Mult}($c_1$, $c_2$, PK):
\begin{itemize}
\item $c_3=c_1.c_2 mod p$
\item return $c_3$
\end{itemize}
\end{itemize}
\item BGV:
\end{itemize}
\subsection{Considered Implementations}

@@

The benchmark problems on which the implementations were evaluated are categorized into three classes : functions operating on bits, functions operating directly on integers or block of bits and functions where branching is required i.e \texttt{if-then-else} condition is evaluated.

Functions operating on bits provide information on the tradeoff for bitwise operations, while when operating directly on integers, the tradeoff is obtained as a function of the size of integers. The first two categories of functions can be evaluated using straight-line programs. For others, evaluation of \texttt{if-then-else} condition on encrypted data is choosen.  The evaluation circuit in this case is longer than the one on non-encrypted data. Hence, this provides information on increase in size of the circuit(when passing from non-encrypted to enrypted data). The addressed problems are :

1. \textbf{XOR of bits}\\
2. \textbf{Sum of :}  
\newline\noindent
\phantom{x}\hspace{3ex} a) bounded integers 
 b) unbounded integers\\ 
3. \textbf{Majority bit} \\
4. \textbf{Sorting : } \newline\noindent
\phantom{x}\hspace{3ex} a) Insertion sort b) Oddd-Even Merge sort \\
	\phantom{x}\hspace{3ex} 	    c) Bitonic sort

The implementations are evaluated on different sorting algorithms to compare the changes in the evaluation circuits.

\subsection{Theoretical Size of Evaluation Circuit}
As the implementations provide addition and multiplicaion of ciphertexts,  size in terms of \texttt{XOR} and \texttt{AND} gates are provided for each problem.

@@

\textbf{Sorting :}
Sorting is tested on a sorting network where comparator gates are used. Each comparator is designed using the following algorithm.

@@

A comparator gate uses $n$ \texttt{AND} gates, $n$ \texttt{OR} gates and $n$ \texttt{NOT} gates to find the larger of the two bit sequences and then to regenerate the maximum and the minuimum $4n$ \texttt{AND} gates and $2n$ \texttt{NOT} and \texttt{OR} gates i.e. $6n$ \texttt{AND} and \texttt{XOR} gates are required. We note that reconstruction of Max and  Min is not required when working on non-encrypted data. Hence, this increases the size of the transformed program operating on encrypted data.

\textbf{Insertion sort :} When allowing for parallel comparators, bubble sort and insertion sort are identical. Hence $n(n-1)/2$ comparator gates are used for insertion sort.

\textbf{Odd-Even Merge sort :} \autoref{oesort} and \autoref{oemerge} are used to perform Odd-Even merge sort and hence  $n/2(log(n)-1) + 1 $ comparator gates are required.

\textbf{Bitonic Sort :} \autoref{bitsort} and \autoref{bitmerge} are used to perform Bitonic Sort and hence $log(n) · (log(n)+1) / 2 $ gates are required.

@@

@@

Anlayze the results previously obtained and its impact on security and complexity. Comparisions of security-cost tradeoff with other schemes like RSA.
\section{Conclusion}

\bibliographystyle{alpha}
\bibliography{article}

%\begin{comment}
\section{Appendix}

 The programs used for Odd-Even Merge sort and Bitonic sort are provided below :

\begin{figure}[h]
\begin{lstlisting}[label = oesort ]

 OddEvenMergeSort(int lo, int n)
if (n > 1)
        {
            int m = n/2;
            oddEvenMergeSort(lo, m);
            oddEvenMergeSort(lo + m, m);
            oddEvenMerge(lo, n, 1);
        }

\end{lstlisting}
\caption{Odd-Even Merge Sort : sort from index lo to n }
\end{figure}

\begin{figure}
\begin{lstlisting}[label = oemerge ]

oddEvenMerge(int lo, int n, int r)
    {
        int m = r*2;
        if (m < n)
        {
	    //even subsequence ;		
            oddEvenMerge(lo, n, m); 
            //odd subsequence ;
	    oddEvenMerge(lo+r, n, m); 
	    int i = lo+r;   
            for (;i + r < lo + n;){
		compare(i, i + r);
		i += m;
	    }
        }
        else
            compare(lo, lo + r);
    }

\end{lstlisting} 
\caption{Odd-Even Merge Sort : merging function}
\end{figure}

\begin{figure}
\begin{lstlisting}[label=bitsort]

sortup( int m, int n) {//from m to m+n
    if (n == 1) return;
    sortup(m, n/2);
    sortdown(m + n/2, n/2);
    mergeup(m, n/2);
}
 sortdown(int m, int n) {//from m to m+n
    if (n == 1) return;
    sortup(m, n/2);
    sortdown(m + n/2, n/2);
    mergedown(m, n/2);
}
\end{lstlisting}
\caption{Bitonic Sort : sort form m to n}
\end{figure}

\begin{figure}[t]
\begin{lstlisting}[label=bitmerge]
mergeup(int m, int n) {  
    if (n == 0) return;

    for (i = 0 : n)
        // Increasing Comparator
	 compare(m + i,m + i + n);
    mergeup(m, n/2);
    mergeup(m + n,n/2);
}

mergedown( int m,  int n) { 
    if (n == 0) return;
    for (i = 0 : n) 
        // Decreasing Comparator
	compare(m + i, m + i + n); 
    mergedown(m, n/2);
    mergedown(m + n, n/2);
}
\end{lstlisting}
\caption{Bitonic Merge} 
\end{figure} 
%\begin{comment}

% sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\balancecolumns
% That's all folks!
\end{document}
